---
title: "Bayesian Data <br>Analysis<br> for <br>Speech Sciences"
subtitle: "Bayesian inference"
author: "Timo Roettger, Stefano Coretta and Joseph Casillas"
institute: "LabPhon workshop"
date: "2021/07/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["https://learnb4ss.github.io/b4ss-theme/css/b4ss.css", "https://learnb4ss.github.io/b4ss-theme/css/b4ss-fonts.css"]
    lib_dir: libs
    nature:
      beforeInit: ["https://learnb4ss.github.io/b4ss-theme/js/b4ss_xaringan.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r setup, include=FALSE}
options(htmltools.dir.version=F)
knitr::opts_chunk$set(fig.retina=2, cache=F, warning=F, message=F)
polite <- learnB4SS::dataset
```

```{r xaringan-extra-all-the-things, echo=F, eval=F}
xaringanExtra::use_xaringan_extra(
  c("tile_view", "panelset", "editable"
    #"scribble", "search", "webcam"
    )
)
xaringanExtra:::use_freezeframe()
```

```{r, 'helpers', echo=F}
source(here::here("assets", "helpers.R"))
```

```{r, load_refs, echo=FALSE, cache=FALSE, warning=F, message=F}
bib <- ReadBib(here("assets", "b4ss_refs.bib"), check = FALSE)
ui <- "- "
```

IMAGE PLACEHOLDER

???

We have been motivating a framework that emphasizes describing our statistical inferences by quantifying uncertainty

This is in contrast to the null hypothesis significance testing framework that was laid out by Timo yesterday and mentioned briefly today

Now we will turn our attention to making statistical inferences under a Bayesian framework

Currently, there are many methods to choose from and I will describe what we believe to be those that are the most common, though this overview will not be exhaustive

---

# Credible intervals

```{r, credible-intervals-1, echo=F, fig.width=14, fig.align="center"}
ggplot(data = tibble(x = rnorm(10e4, 15, 2.5))) +  
  aes(x = x, fill = stat(x < 10 | x > 20)) + 
  stat_halfeye(show.legend = F, size = -10, point_size = -10) +
  scale_fill_manual(values = c(b4ss_colors[[1]], b4ss_colors[[2]])) + 
  theme_void() + 
  NULL
```

???

If you have heard of Bayesian inference before, there is a good chance you have also heard of credible intervals

We could describe them as the Bayesian counterpart to confidence intervals under a frequentist framework. 

Some might say the difference is merely philosophical, but I won't go into that now. 

Crucially, credible intervals are what most people think confidence intervals represent: that is, the interval within which an observed parameter estimate falls with a particular probability

When we describe a posterior distribution we typically say that a specified % of the posterior falls within a certain range

In this plot I am using color highlight part of the posterior that falls within a certain range

---
class: middle

# Credible intervals

```{r, credible-intervals-2, echo=F, fig.width=14, fig.align="center"}
ggplot(data = tibble(x = rnorm(10e4, 15, 2.5))) +  
  aes(x = x, fill = stat(x < 10 | x > 20)) + 
  stat_halfeye(show.legend = F, size = -10, point_size = -10) +
  geom_segment(
    data = tibble(y = 0, yend = 0.9, x = c(10, 20), xend = c(10, 20)), 
    aes(y = y, yend = yend, x = x, xend = xend), lty = 2, size = 1) + 
  scale_fill_manual(values = c(b4ss_colors[[1]], b4ss_colors[[2]])) + 
  geom_text(
    data = tibble(
      x = c(10, 15, 20, 15), 
      y = c(-0.025, -0.06, -0.025, 1), 
      label = c("10", expression(beta), "20", expression(paste("95% credible interval")))
      ), 
    aes(x = x, y = y, label = label), color = "black", size = 8, parse = T) + 
  theme_void() + 
  NULL
```

???

To give a more concrete example, you might hear somebody say something like: the probability that B ("a value") is between 10 and 20 is .95 (or "we're 95% sure that the value is between 10 and 20")

This is an informative way to quantify uncertainty when describing a posterior distribution

This is the case with the plot you see here... we're 95% certain that the value of beta falls between 10 and 20

---
class: middle

# Credible intervals

```{r, credible-intervals-3, echo=F, fig.width=14, fig.align="center"}
ggplot(data = tibble(x = rnorm(10e4, 15, 2.5))) + 
  aes(x = x, fill = stat(x < 10 | x > 20)) + 
  stat_halfeye(show.legend = F, size = -10, point_size = -10) +
  geom_vline(xintercept = 0, lty = 3, size = 1) + 
  scale_fill_manual(values = c(b4ss_colors[[1]], "grey80")) + 
  labs(y = NULL, x = expression(beta)) + 
  b4ss_bw(base_size = 20) + 
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
    panel.border = element_blank(), 
    axis.line.x = element_line(color = "black", size = 0.25)) + 
  NULL
```

???

Some researchers use a 95% CI to make statistical inferences by saying something like "if the 95% credible interval of the posterior distribution for a given parameter does not include 0, then we will consider that to be compelling evidence for an effect"

That would certainly be the case in the plot we see here, as the entire posterior falls far from 0, but obviously this is not always the case

---
class: middle

# Credible intervals

```{r, credible-intervals-4, echo=F, fig.width=14, fig.align="center"}
ggplot(data = tibble(x = rnorm(10e4, 3, 2.5))) + 
  aes(x = x, fill = stat(x < -2 | x > 8)) + 
  stat_halfeye(show.legend = F, size = -10, point_size = -10) +
  geom_vline(xintercept = 0, lty = 3, size = 1) + 
  scale_fill_manual(values = c(b4ss_colors[[1]], "grey80")) + 
  labs(y = NULL, x = expression(beta)) + 
  b4ss_bw(base_size = 20) + 
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
    panel.border = element_blank(), 
    axis.line.x = element_line(color = "black", size = 0.25)) + 
  NULL
```

???

In this plot, for example, the 95% CI of the posterior overlaps 0 by a pretty large amount

Thus some researchers would conclude that there is not compelling evidence for this particular effect

Notice that I've used 95% for the range here and in the previous examples 

This is without a doubt the most common value, which is carryover from NHST (p < 0.05) and completely arbitrary 

You can use any interval you want to describe the posterior

Let's practice a couple examples

---
class: title-slide-section-blue, middle, center

# Live coding

<!--
short practical
-->

<!--
CI Practical: Describe posteriors as a way to do inference
USE MODEL FROM STEFs EXAMPLE FROM PREVIOUS DAY
-->

```{r, credible-intervals, include=F}
# Fake posterior (for ex.)
posterior <- rnorm(n = 1000, mean = 5, sd = 3)

# Posterior from int only model (until I get stefs model)
ar_bayes_int <- brm(articulation_rate ~ 1, data = polite, 
  file = here("assets", "ar_int"))

# base r example of getting quantile of distribution
quantile(posterior, c(0.025, 0.975))

# bayesTestR to simplify
hdi_ex <- hdi(posterior)
hdi_ex
plot(hdi_ex)

# Example with model
# result <- model_parameters(ar_bayes_int, effects = "all", components = "all")
# plot(result, show_intercept = T, n_columns = NULL)

```

---





class: title-slide-section-grey, middle

# .RUred[Probability of direction]

IMAGE PLACEHOLDER

---

# Probability of direction

IMAGE PLACEHOLDER

???

Another option decision making is to use the probability of direction, also known as the maximum probability of effect

This value can range from 50% to 100% (0.5 - 1.0)

It is the proportion of the posterior that has the same sign (that is, if its + or -) as the median of the distribution

In simple terms, it tells us the probability that a given parameter estimate is positive or negative

To illustrate, if you have a probability of direction of 100%, then you are 100% certain that a given parameter is positive (or negative)

On the other hand, if you obtain a probability of direction of 50%, then the parameter is equally likely to be positive as it is to be negative (that is a lot of uncertainty!)

The pbe is likely the closest equivalent to a p-value

It is quite simple to calculate, so let's look at a few examples...

---
class: title-slide-section-blue, middle, center

# Live coding

<!--
short practical
-->

```{r, include=F}
test <- p_direction(ar_bayes_int, component = "all")
plot(test, show_intercept = T, n_columns = NULL, priors = T)
```











---
class: title-slide-section-grey, middle

# .RUred[Region of practical <br>equivalence] 

### (ROPE)

IMAGE PLACEHOLDER

---

# ROPE

IMAGE PLACEHOLDER

???



---
class: title-slide-section-blue, middle, center

# Live coding

<!--
short practical
-->

```{r, include=F}
test <- rope(ar_bayes_int, ci = c(0.9, 0.95), effects = "all", component = "all")
plot(test, show_intercept = T, n_columns = NULL, rope_color = "white")
```








---
class: title-slide-section-grey, middle

# Other methods

---

# Define your hypothesis

IMAGE PLACEHOLDER

???

Brief explanation and back to R

---
class: title-slide-section-blue, middle, center

# Live coding

<!--
SHORT PRACTICAL
brms::hypothesis()

- minor wrangling posteriors (ex. pr that B is > X value)
- testing hypotheses (pr that B is > X value but using hypothesis function)
-->


---

# Bayes factors

IMAGE PLACEHOLDER

???

- Bayes factors (only conceptually)

---

# Summary

- Credible intervals

- Probability of direction (maximum probability of effect)

- Region of practical equivalence (ROPE)

- User defined hypotheses

- Bayes factors

???

We have touched on some of the many ways in which one can use the posterior for statistical inferences

The main focus has been on emphasizing different ways we can use these tools to quantify uncertainty and help us in making decisions

They provide a rich set of tools at the researchers disposal that contrasts with the sole p-value available under a frequentist/NHST framework

---
class: title-slide-section-blue, middle, center

# Live coding

<!--
PRACTICAL (longer): Run model on Bodo’s data with prepared RMarkdown

think of new hypotheses to test, use your favorite method (or all of them)
--> 
